---
title: "100s of Regression Models with 10 lines of R code"
subtitle: "With purrr and gapminder package"
author: "FRK"
date: "6/17/2021"
output: 
  html_document:
    theme: united
    toc: TRUE
    highlight: tango
---

```{r, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


# Problem

Today we are going to fit a simple regression model with yearly population to year, by each country of the world. The data is available on the *gapminder* package. And we will use *functional programming* to apply regression model to each country with an easy 10 lines of code.

# Plot the data by Country and Continent

```{r, fig.width = 10}
require(tidyverse)
require(gapminder)
head(gapminder)

gapminder %>% 
  ggplot(aes(year, pop, color = country)) +
  geom_line(show.legend = F) +
  facet_wrap(~continent, scales = "free_y")

```

We can clearly see some difference within and between continents. There is also a linear pattern in increase of population with year. Lets fit the regression model and check which countries and continent have highest yearly population growth rate.


# Fit the Regression models by each country

We are going to use the *map* function from the *purrr* package to apply regression model to each country.

```{r}
require(broom)
require(purrr)

reg_models <- gapminder %>%  
  group_by(country, continent) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(pop ~ year, data = .x))) %>% 
  mutate(coefs = map(models, ~tidy(.x))) %>% 
  unnest(coefs) 

### FItted Models and Coefficients
reg_models     
   
```

# Plot the Estimated Growth Rate

```{r, fig.width = 10}

g1 <- reg_models %>% 
  filter(term == "year") %>% 
  arrange(estimate) %>% 
  unnest(data) %>% 
  group_by(country, continent, estimate) %>% 
  nest() %>% 
  ggplot(aes(x = reorder(country, estimate), y = estimate, fill = continent)) +
  geom_col() + 
  scale_y_log10() +
  theme_light() +
  theme(axis.text.x = element_text(size=3, angle = 90),
        panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
  labs(title = "Estimated per year Population Growth Rate by Country",
       x = "Country",
       y = "Coefficient / Estimate on log_10 scale",
       fill = "Continent")
 
plotly::ggplotly(g1)
        
```

Looks like Asia, Africa and South American countries have the highest population growth rates.

# Digging Deep to what's happening with the code

```{r}

### First Group by Country and COntinent
gapminder %>%  
  group_by(country, continent)

### The nest the data
gapminder %>%  
  group_by(country, continent) %>% 
  nest() 

### Apply Linear Regression to the column data
### Inside the data, there are columns called pop and year
### .x means take everything from each tibble of the data
gapminder %>%  
  group_by(country, continent) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(pop ~ year, data = .x)))

### We created the model column
### Now we tidy the model column
### This created tibbles of coefficients foe each country
gapminder %>%  
  group_by(country, continent) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(pop ~ year, data = .x))) %>% 
  mutate(coefs = map(models, ~tidy(.x)))

### Finally unnest the coefs column to get our estimates
gapminder %>%  
  group_by(country, continent) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(pop ~ year, data = .x))) %>% 
  mutate(coefs = map(models, ~tidy(.x))) %>% 
  unnest(coefs) 
```













