---
title: "100s of Regression Models with 10 lines of R code"
subtitle: "With purrr and gapminder package"
author: "FRK"
date: "6/17/2021"
output: 
  html_document:
    theme: united
    toc: TRUE
    highlight: tango
---

```{r, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


# Problem

Today we are going to fit a simple regression model with yearly population to year, by each country of the world. The data is available on the *gapminder* package. And we will use *functional programming* to apply regression model to each country with an easy 10 lines of code.

# Plot the data by Country and Continent

```{r, fig.width = 10}
require(tidyverse)
require(gapminder)
head(gapminder)

gapminder %>% 
  ggplot(aes(year, pop, color = country)) +
  geom_line(show.legend = F) +
  facet_wrap(~continent, scales = "free_y") +
  scale_y_log10() # Took log to make the y axis readable

```

We can clearly see some difference within and between continents. There is also a linear pattern in increase of population with year. Lets fit the regression model and check which countries and continent have highest yearly population growth rate.


# Fit the Regression models by each country

- We are going to use the *map* function from the *purrr* package to apply regression model to each country.
- *tidy* makes model output tidy tibble
- *glance* creates tibble of model's Goodness of Fit estimates

```{r}
require(broom)
require(purrr)

reg_models <- gapminder %>%  
  group_by(country, continent) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(pop ~ year, data = .x))) %>% 
  mutate(coefs = map(models, ~tidy(.x)),
         GOF = map(models, ~glance(.x))) %>% 
  unnest(coefs) 

### FItted Models and Coefficients
reg_models     
   
```

# Plot the Estimated Growth Rate

```{r, fig.width = 10}

g1 <- reg_models %>% 
  filter(term == "year") %>% 
  arrange(estimate) %>% 
  unnest(data) %>% 
  group_by(country, continent, estimate) %>% 
  nest() %>% 
  ggplot(aes(x = reorder(country, estimate), y = estimate, fill = continent)) +
  geom_col() + 
  scale_y_log10() +
  theme_light() +
  theme(axis.text.x = element_text(size=3, angle = 90),
        panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()) +
  labs(title = "Estimated per year Population Growth Rate by Country",
       x = "Country",
       y = "Coefficient / Estimate on log_10 scale",
       fill = "Continent")
 
plotly::ggplotly(g1)
        
```

Looks like Asia, Africa and South American countries have the highest population growth rates.

# Pvalue, Estimate, R-Squared in log scale

- Taking log scale to make the plot readable.
- Large Inverse-R-Squared states poor fit to linear model.
- Point above horizontal line indicate Insignificant p-value for the estimate. That is estimate is not significantly greater then zero.
- Larger coefficient value means higher population growth rate. Countries on right side have higher growth rate.

```{r, fig.width = 12, fig.height=12}
require(ggrepel)
reg_models %>% 
  filter(term == "year") %>% 
  select(-std.error, -statistic, -p.value) %>% 
  unnest(GOF) %>% 
  mutate(`Inverse-R-Squared` =  1/ round(r.squared, 2)) %>% 
  
  ggplot(aes(x = estimate, y = p.value, color = continent)) +
  geom_point(aes(size = `Inverse-R-Squared`)) +
  
  geom_hline(yintercept = 0.05, 
             linetype = "longdash", 
             alpha = 0.6) +
  
  labs(shape="Inverse R Squared", colour="Continents",
       x = "Coefficient Value",
       y = "P-value of Estimate in Log10 Scale", 
       title = "Coefficient, Pvalue of Coefficients, R-Squared",
       caption = "Both Axis are free, so not directly comparable continent wise\nY axis on log scale, X axis is abosolute") +
  
  scale_y_log10() +
  scale_x_log10() +
  
  geom_text_repel(aes(label = country), 
                   size = 2.2, 
                   alpha = 0.5, 
                   show.legend = F, 
                   max.overlaps = 20, 
                   min.segment.length = unit(0, 'lines')) +
  
  facet_wrap(~continent, scales = "free")  +
  theme_light() +
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",  
        legend.box = "vertical")
```

***

# Digging Deep to what's happening with the code

```{r}

### First Group by Country and COntinent
gapminder %>%  
  group_by(country, continent)

### The nest the data
gapminder %>%  
  group_by(country, continent) %>% 
  nest() 

### Apply Linear Regression to the column data
### Inside the data, there are columns called pop and year
### .x means take everything from each tibble of the data
gapminder %>%  
  group_by(country, continent) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(pop ~ year, data = .x)))

### We created the model column
### Now we tidy the model column
### This created tibbles of coefficients foe each country
gapminder %>%  
  group_by(country, continent) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(pop ~ year, data = .x))) %>% 
  mutate(coefs = map(models, ~tidy(.x)))

### Finally unnest the coefs column to get our estimates
gapminder %>%  
  group_by(country, continent) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(pop ~ year, data = .x))) %>% 
  mutate(coefs = map(models, ~tidy(.x))) %>% 
  unnest(coefs) 

### ALso unnest the GOF column to get our GOF estimates
gapminder %>%  
  group_by(country, continent) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(pop ~ year, data = .x))) %>% 
  mutate(coefs = map(models, ~tidy(.x)),
         GOF = map(models, ~glance(.x))) %>% 
  unnest(GOF) 
```













